\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2019-02-15 Documents/M1/Astronum_2018/sections/Acknowledge.tex

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725. This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research. This research was also supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration, and by the National Science Foundation Gravitational Physics Program (NSF-GP 1505933 and 1806692).

>>>>

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U.S. Department of Energy under Contract No. De-AC05-00OR22725. This material is based, in part, on work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research. This research was also supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration, and by the National Science Foundation Gravitational Physics Program (NSF-GP 1806692).

### 2019-02-15 Documents/M1/Astronum_2018/sections/Conclusion.tex

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
We also ran our code with Minerbo closure, IMEX PC2, IMEX SSP2332 and IMEXRKCB2 schemes.
And the results show that only PD-ARS schemes have stability.
In addition, closures have impact on the simulation result.
Even though RKCB2 with Minerbo closure luckily survived in our test, the result it gives is compromised.
In what way and what degree the results are in fact compromised either by the closure or by a particular correction step for unrealizable moments are difficult to determine and left for further study.

>>>>

In the applications that motivate this work, the neutrino distribution function can vary from 0 to 1.  
Hence, we have considered algebraic closures based on Fermi-Dirac statistics for both low and high occupancy.  
Among the seven algebraic closures we considered -- Kershaw~\cite{kershaw_1976}, Wilson~\cite{wilson_1975,leblancWilson_1970}, Levermore~\cite{levermore_1984}, Minerbo~\cite{minerbo_1978}, Janka 1~\cite{janka_1991}, Janka 2~\cite{janka_1992}, and Cernohorsky \& Bludman~\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure obeys Fermi-Dirac statistics for all occupancies.  
As a result, we employed the Cernohorsky \& Bludman closure for the neutrino stationary state test in Section~\ref{se: Neutrino Stationary State Test}.
We also ran our code with Minerbo closure, IMEX PC2, IMEX SSP2332 and IMEXRKCB2 schemes, and the results show that only PD-ARS schemes have stability.
In addition, closures have impact on the simulation result.
As we observed, both using PD-ARS scheme, there are $\sim30\%$ difference in the neutrino number densities (and relaxation time) between the results obtained with Minerbo closure and that with CB closure.
Even though RKCB2 with Minerbo closure luckily survived our test, the results it gave were compromised: closer to PD-ARS scheme with Minerbo closure other than PD-ARS scheme with CB closure.
In what way and to what degree the results are in fact compromised either by the closure or by a particular correction step for unrealizable moments are difficult to determine fully and is left for further study.

### 2019-02-15 Documents/M1/Astronum_2018/sections/Numerical.tex

The sine wave streaming test is designed to test accuracy in the free-steaming regime; i.e. $\sigma_{\Ab} = \sigma_{\Scatt} = 0$.
A periodic domain of unit length is used and the initial condition is $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$.
We evolve the test until the sine wave has completed 10 crossings of the domain.
Figure~\ref{fig: SineWaveStreaming} plots the absolute error for the number density versus the number of elements $N$.
We see the errors obtained with SSPRK3 and PD-ARS3 are smallest and decrease as $N^{-3}$, as expected for a scheme combining third-order accurate time stepping with third-order accurate spatial discretization.
For all the other schemes, using second-order accurate explicit time stepping, the error decreases as $N^{-2}$.
Among the second-order schemes, SSP2332 has the smallest error.
In the streaming limit, the PD-ARS schemes reduce to SSPRK schemes, such that PD-ARS2 to SSPRK2 and PD-ARS3 to SSPRK3, respectively.
Therefore, the absolute errors of PD-ARS schemes and SSPRK schemes are indistinguishable.

@@

The next test, adapted from~\cite{skinnerOstriker_2013}, is designed for absorption-dominated regimes, with $\sigma_{\Scatt} = 0$ and $f_0 = 0$, which results in exponential damping of the wave amplitude.
A periodic domain $D=\{x:x\in[0,1]\}$ and initial condition $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$ are used.
The amplitude of the analytical solution decreases as $e^{-\sigma_{\Ab} t}$.
For $\sigma_{\Ab}$ = 0.1, 1 and 10 we evolve the test until the initial condition has been damped by a factor $e^{-10}$. 
Figure~\ref{fig:SineWaveDamping} shows convergence results of the test in the relative error.
Results for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.
SSP2332 is the most accurate among these schemes for $\sigma_{\Ab}= 1$ and 10. 
For $\sigma_{\Ab}=0.1$, PD-ARS2 is the most accurate scheme for $N=8$ and $N=16$. 
We have seen the same behavior for the scheme proposed by McClarren et al.~\cite{mcclarren_etal_2008} (PC2) in~\cite{chu_etal_2018}.
Since these are special cases for $N=8$ and $N=16$, we won't recommend PD-ARS2 instead of SSP2332 for the damping problem.
Only SSP2332, a second-order accurate scheme, displays a second-order convergence rate.  
The PD-ARS schemes are first-order accurate.

@@

The final test with known smooth solution, adopted from~\cite{radice_etal_2013}, is the sine wave diffusion test; i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain $D=\{x:x\in[-3,3]\}$ with initial conditions $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ are used.  
The reference diffusion solution is given by $\cJ = \cJ_{0} \times \exp\big(-\f{\pi^2 t}{27\sigma_{\Scatt}}\big)$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.  
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$, 
at which time the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the absolute error, obtained using different values of $\sigma_{\Scatt}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$, versus $N$. 
Results for $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
For $\sigma=10^{2}$, the errors in the number density $\cJ$ do not drop below $10^{-6}$ because of differences between the two-moment model and the diffusion equation used to obtain the analytic solution.
For larger values of the scattering opacity, $\sigma=10^{3}$ or $10^{4}$, the two-moment model agrees better with the diffusion model, and we observe convergence over the entire range of $N$.
PD-ARS2 behaves as well as SSP2332 in the diffusion region but requires 33\% less implicit solves per time step.

@@

In this section we consider a more ``realistic'' test: two-dimensional multigroup neutrino transport with emission, absorption, and isoenergetic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from~\cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction, a reflecting inner boundary and a homogeneous out-flowing outer boundary.
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed both CB and Minerbo closures.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from~\cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley~\cite{cavaglieriBewley2015}, and the IMEX PC2 scheme proposed by McClarren et al.~\cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state with either CB or Minerbo closure.
SSP2332 and IMEX PC2 failed after a few time step with neither CB or Minerbo closure because of unrealizable moment.
Even though IMEXRKCB2 with Minerbo closure can run and reach a steady state in this test, its results are different from that of PD-ARS schemes with CB closure and there is no guarantee for its stability.
%It's difficult to determine whether the results are in fact compromised by a particular correction step and, if so, in what way and to what degree.

Results obtained with the IMEX PD-ARS schemes are plotted in Figure~\ref{fig:NeutrinoStationaryTestEvolve} for various times: $t=0.01$~ms (top panels), 0.35~ms (middle panels), and 5.0~ms (bottom panels).  
In the left column we plot the solution in the $|\vect{\cH}|\cJ$-plane.  
In the middle column we show scatter plots of the number density $\cJ$ versus radius for select neutrino energies: 5~MeV (red lines), 16~MeV (magenta lines), and $93$~MeV (blue lines).  
In the right column we plot the flux factor $|\vect{\cH}|/\cJ$ versus radius for the same neutrino energies as in the middle column. 
In the left panels, each solution point in the domain is marked as a red dot in the $|\vect{\cH}|\cJ$-plane, and the realizable domain is shown as the light blue region.  
The figures show that all the states in the simulation with the PD-ARS schemes are realizable.
In the middle and right column, we can see how the neutrinos are generated near the core, stream out, and reach an equilibrium distribution over the phase space eventually.
The oscillation in the right column is caused by numerical error.
%can see how the scattering reshapes the flux direction. 
%The oscillatory behavior, which is more obvious for high-energy groups, is due to numerical error.
%by the oscillatory in $|\bcH|$ as a result of scattering: scattering changes neutrino's direction and brings randomness to $|\bcH|$.  
%High-energy groups can gain more randomness due to machine accuracy error.

>>>>

The sine wave streaming test is designed to test accuracy in the free-steaming regime; i.e. $\sigma_{\Ab} = \sigma_{\Scatt} = 0$.
A periodic domain of unit length is used and the initial condition is $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$.
We evolve the test until the sine wave has completed 10 crossings of the domain.
Figure~\ref{fig: SineWaveStreaming} plots the absolute error for the number density versus the number of elements $N$.
We see the errors obtained with SSPRK3 and PD-ARS3 are smallest and decrease as $N^{-3}$, as expected for a scheme combining third-order accurate time stepping with third-order accurate spatial discretization.
For all the other schemes, using second-order accurate explicit time stepping, the error decreases as $N^{-2}$.
Among the second-order schemes, SSP2332 has the smallest error.
In the streaming limit, the PD-ARS schemes reduce to SSPRK schemes --- PD-ARS2 to SSPRK2 and PD-ARS3 to SSPRK3, respectively.
Therefore, the absolute errors of PD-ARS schemes and SSPRK schemes are indistinguishable.

@@

The next test, adapted from~\cite{skinnerOstriker_2013}, is designed for absorption-dominated regimes, with $\sigma_{\Scatt} = 0$ and $f_0 = 0$, which results in exponential damping of the wave amplitude.
A periodic domain $D=\{x:x\in[0,1]\}$ and initial condition $\cJ_{0} = \bcH_{0} = 0.5+0.49\times\sin\big(2\pi\,x\big)$ are used.
The amplitude of the analytical solution decreases as $e^{-\sigma_{\Ab} t}$.
For $\sigma_{\Ab}$ = 0.1, 1 and 10 we evolve the test until the initial condition has been damped by a factor $e^{-10}$. 
Figure~\ref{fig:SineWaveDamping} shows convergence results of the test in the relative error.
Results for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.
SSP2332 is the most accurate among these schemes for $\sigma_{\Ab}= 1$ and 10. 
For $\sigma_{\Ab}=0.1$, PD-ARS2 is the most accurate scheme for $N=8$ and $N=16$. 
We have seen the same behavior for the scheme proposed by McClarren et al.~\cite{mcclarren_etal_2008} (PC2 in~\cite{chu_etal_2018}).
Since these are special cases for $N=8$ and $N=16$, we do not recommend PD-ARS2 instead of SSP2332 for the damping problem.
Only SSP2332, a second-order accurate scheme, displays a second-order convergence rate.  
The PD-ARS schemes are first-order accurate.

@@

The final test with known smooth solutions, adopted from~\cite{radice_etal_2013}, is the sine wave diffusion test; i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain $D=\{x:x\in[-3,3]\}$ with initial conditions $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ are used.  
The reference diffusion solution is given by $\cJ = \cJ_{0} \times \exp\big(-\f{\pi^2 t}{27\sigma_{\Scatt}}\big)$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.  
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$, 
at which time the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the absolute error, obtained using different values of $\sigma_{\Scatt}$, for various IMEX schemes at $t=t_{\mbox{\tiny end}}$, versus $N$. 
Results for $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
For $\sigma=10^{2}$, the errors in the number density $\cJ$ do not drop below $10^{-6}$ because of differences between the two-moment model and the diffusion equation used to obtain the analytic solution.
For larger values of the scattering opacity, $\sigma=10^{3}$ or $10^{4}$, the two-moment model agrees better with the diffusion model, and we observe convergence over the entire range of $N$.
PD-ARS2 behaves as well as SSP2332 in the diffusion region but requires 33\% less implicit solves per time step.

@@

In this section we consider a more ``realistic'' test: two-dimensional multigroup neutrino transport with emission, absorption, and isoenergetic scattering through a stationary background.  
This test is designed to test the realizability-preserving properties of the PD-ARS schemes.  
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities used (adopted from~\cite{Bruenn_1985}).  
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction, a reflecting inner boundary and a homogeneous out-flowing outer boundary.
We use 10 energy groups to cover the range from 0~MeV to 300~MeV and initialize the neutrino number density to $\cJ = 10^{-99}$, flux density to $\bcH=0$, and evolve until an approximate steady state is reached ($t=5$~ms).  
For this test, we employed both CB and Minerbo closures.  
We attempted to run this test with our PD-ARS schemes, SSP2332 from~\cite{pareschiRusso_2005}, IMEXRKCB2 proposed by Cavaglieri \& Bewley~\cite{cavaglieriBewley2015}, and the IMEX PC2 scheme proposed by McClarren et al.~\cite{mcclarren_etal_2008}.
Only the PD-ARS schemes produce realizable moments and are able to evolve to a steady state with either CB or Minerbo closure.
SSP2332 and IMEX PC2 failed after a few time steps with either CB or Minerbo closure because of the development of unrealizable moments.
Even though IMEXRKCB2 with Minerbo closure can run and reach a steady state in this test, its results are different from that of PD-ARS schemes with CB closure, and there is no guarantee of stability.
%It's difficult to determine whether the results are in fact compromised by a particular correction step and, if so, in what way and to what degree.

Results obtained with the IMEX PD-ARS schemes are plotted in Figure~\ref{fig:NeutrinoStationaryTestEvolve} for various times: $t=0.01$~ms (top panels), 0.35~ms (middle panels), and 5.0~ms (bottom panels).  
In the left column we plot the solution in the $|\vect{\cH}|\cJ$-plane.  
In the middle column we show scatter plots of the number density $\cJ$ versus radius for select neutrino energies: 5~MeV (red lines), 16~MeV (magenta lines), and $93$~MeV (blue lines).  
In the right column we plot the flux factor $|\vect{\cH}|/\cJ$ versus radius for the same neutrino energies as in the middle column. 
In the left panels, each solution point in the domain is marked as a red dot in the $|\vect{\cH}|\cJ$-plane, and the realizable domain is shown as the light blue region.  
The figures show that all the states in the simulation with the PD-ARS schemes are realizable.
In the middle and right column, we can see how the neutrinos are generated near the core, stream out, and eventually reach an equilibrium distribution over the phase space.
The oscillation in the right column is caused by numerical error.
%can see how the scattering reshapes the flux direction. 
%The oscillatory behavior, which is more obvious for high-energy groups, is due to numerical error.
%by the oscillatory in $|\bcH|$ as a result of scattering: scattering changes neutrino's direction and brings randomness to $|\bcH|$.  
%High-energy groups can gain more randomness due to machine accuracy error.

### 2019-02-15 Documents/M1/Astronum_2018/sections/intro.tex

One essential part of the explosion mechanism is neutrino transport.
Neutrino energy deposition is believed to be the major driver of CCSN explosions, -- except in peculiar cases where rapid rotation is present and magnetohydrodynamic effects may dominate (for reviews, see~\cite{mezzacappa_2005,janka_2012,burrows_2013,muller_2016}).
Ideally, neutrino transport would be modeled by the Boltzmann transport equation, which is an integro-partial-differential equation evolving a phase-space distribution function $f$.~(e.g., see~\cite{mezzacappaBruenn_1993a,mezzacappaBruenn_1993b,mezzacappaBruenn_1993c,mezzacappa_etal_2001,liebendoerfer_etal_2001,liebendoerfer_etal_2004,livne_etal_2004,liebendoerfer_etal_2005,ott_etal_2008,sumiyoshiYamada_2012,nagakura_etal_2014,sumiyoshi_etal_2014,nagakura_etal_2018} for studies of CCSN with Boltzmann transport in various approximate settings.)
Simulating the neutrino transport implies finding a solution of the Boltzmann equation for a specific domain and period, with acceptable accuracy.

However, solving the Boltzmann transport equation with sufficient phase-space resolution and full weak interaction physics is at present too expensive.
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted (e.g., see~\cite{kuroda_etal_2016,roberts_etal_2016,just_etal_2018,vartanyan_etal_2018}).
Using the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$ -- the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the equation of $\bcH$ includes the second angular moment $\bcK$.
Knowledge of $\bcK$ is needed to close the two-moment system.
Therefore, a closure that gives $\bcK$ consistent with $\cJ$ and $\bcH$ is needed.
The better the closure predicts $\bcK$, the more accurate the two-moment method. 
The two-moment method has been widely applied in the CCSN modeling community with different algebraic closures, such as the Minerbo~\cite{minerbo_1978} closure (e.g.~{O'Connor} and {Couch}~\cite{oConnorCouch_2018}, Pan and et al.~\cite{pan_etal_2018}, Glas et al.~\cite{glas_etal_2018}, and Just et al.~\cite{just_etal_2018}) and the Levermore~\cite{levermore_1984} closure (e.g.~Vartanyan et al.~\cite{vartanyan_etal_2018}, Cabezon et al.~\cite{cabezon_etal_2018}, and Kuroda et al.~\cite{kuroda_etal_2016}).

Applying the two-moment method does simplify the problem, but doesn't guarantee an affordable solution.
To be precise, how to discretize the continuous equation of system given by the two-moment method and solve the discretized system efficiently remains a question.
In fact, the time scales of neutrino interactions with the background (can be $\sim\mathcal{O}(10^{-13})$~second) is short compared to the duration of the CCSN explosion ($\sim\mathcal{O}(1)$~second).  
This means that $\sim\mathcal{O}(10^{13})$ time steps would be needed for solving the system fully explicitly. 
On the other hand, solving the moment equations fully implicitly requires inverting global band-structured matrices whose sizes depend on the phase-space discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require far fewer time steps compared with a fully explicit method, and the computation for each step is easily parallelizable.
For the relativistic setting that we have, where the fluid and the neutrinos have comparable propagation speeds, and IMEX methods can be efficient.

To model neutrino transport using a two-moment method, two (or at least two) things need to be chosen carefully: an algebraic closure based on Fermi-Dirac statistics for closing the two-moment equations and a convex-invariant, diffusion-accurate IMEX scheme to ensure a physical result.
A convex-invariant scheme has the following property: if the solution $u^{n}\in W$ and $W$ is convex, then $u^{n+1}\in W$.
Since the neutrino distribution function is bounded ($f\in[0,1]$) by the Pauli exclusion principle, its moments as weighted integrals of a bounded function over the domain $\omega\in\bbS^{2}$ are also bounded.
We call the moments satisfying the constraints due to Fermi-Dirac statistics \textit{realizable moments} and the realizable $\cJ$ and $\bcH$ define a convex set~\cite{chu_etal_2018}.
The algebraic closure should give a realizable $\bcK$, and the well-posedness of the closure requires realizable $\cJ$ and $\bcH$.
It explains why an algebraic closure based on Fermi-Dirac statistics is needed.
Realizability of $\cJ$ and $\bcH$ after each time step requires a convex-invariant IMEX scheme.
Since the realizable moments form a convex set, it is possible to construct a realizability-preserving method using a convex-invariant IMEX scheme for two-moment neutrino transport.
In addition, the physics of neutrino transport in CCSNe requires the IMEX scheme to be diffusion-accurate.

>>>>

One essential part of the explosion mechanism is neutrino transport.
Neutrino energy deposition is believed to be the major driver of CCSN explosions, except in peculiar cases where rapid rotation is present and magnetohydrodynamic effects may dominate (for reviews, see~\cite{mezzacappa_2005,janka_2012,burrows_2013,muller_2016}).
Ideally, neutrino transport would be modeled by the Boltzmann transport equation, which is an integro-partial-differential equation evolving a phase-space distribution function $f$~(e.g., see~\cite{mezzacappaBruenn_1993a,mezzacappaBruenn_1993b,mezzacappaBruenn_1993c,mezzacappa_etal_2001,liebendoerfer_etal_2001,liebendoerfer_etal_2004,livne_etal_2004,liebendoerfer_etal_2005,ott_etal_2008,sumiyoshiYamada_2012,nagakura_etal_2014,sumiyoshi_etal_2014,nagakura_etal_2018} for studies of CCSN with Boltzmann transport in various approximate settings).
Simulating the neutrino transport implies finding a solution of the Boltzmann equation for a specific domain and period, with acceptable accuracy.

However, solving the Boltzmann transport equation with sufficient phase-space resolution and full weak interaction physics is at present too expensive.
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted (e.g., see~\cite{kuroda_etal_2016,roberts_etal_2016,just_etal_2018,vartanyan_etal_2018}).
Using the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$ -- the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the equation of $\bcH$ includes the second angular moment $\bcK$.
Knowledge of $\bcK$ is needed to close the two-moment system.
Therefore, a closure that gives $\bcK$ consistent with $\cJ$ and $\bcH$ is needed.
The better the closure predicts $\bcK$, the more accurate the two-moment method will be. 
The two-moment method has been widely applied in the CCSN modeling community with different algebraic closures, such as the Minerbo~\cite{minerbo_1978} closure (e.g.~{O'Connor} and {Couch}~\cite{oConnorCouch_2018}, Pan and et al.~\cite{pan_etal_2018}, Glas et al.~\cite{glas_etal_2018}, and Just et al.~\cite{just_etal_2018}) and the Levermore~\cite{levermore_1984} closure (e.g.~Vartanyan et al.~\cite{vartanyan_etal_2018}, Cabezon et al.~\cite{cabezon_etal_2018}, and Kuroda et al.~\cite{kuroda_etal_2016}).

Applying the two-moment method does simplify the problem, but doesn't guarantee an affordable solution.
To be precise, how to discretize the continuous system of equation given by the two-moment method and solve the discretized system efficiently remains a question.
In fact, the time scales of neutrino interactions with the background (they can be $\sim\mathcal{O}(10^{-13})$~second) is short compared to the duration of the CCSN explosion ($\sim\mathcal{O}(1)$~second).  
This means that $\sim\mathcal{O}(10^{13})$ time steps could be needed for solving the system fully explicitly. 
On the other hand, solving the moment equations fully implicitly requires inverting global band-structured matrices whose sizes depend on the phase-space discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require far fewer time steps compared with a fully explicit method, and the computation for each step is easily parallelizable.
For the relativistic setting that we have, where the fluid and the neutrinos have comparable propagation speeds, IMEX methods can be efficient.

To model neutrino transport using a two-moment method, two (or at least two) things need to be chosen carefully: an algebraic closure based on Fermi-Dirac statistics for closing the two-moment equations and a convex-invariant, diffusion-accurate IMEX scheme to ensure a physical result.
A convex-invariant scheme has the following property: if the solution $u^{n}\in W$ and $W$ is a convex set, then $u^{n+1}\in W$.
Since the neutrino distribution function is bounded ($f\in[0,1]$) by the Pauli exclusion principle, its moments as weighted integrals of a bounded function over the domain $\omega\in\bbS^{2}$ are also bounded.
We call the moments satisfying the constraints due to Fermi-Dirac statistics \textit{realizable moments} and the realizable $\cJ$ and $\bcH$ define a convex set~\cite{chu_etal_2018}.
The algebraic closure should give a realizable $\bcK$, and the well-posedness of the closure requires realizable $\cJ$ and $\bcH$.
This explains why an algebraic closure based on Fermi-Dirac statistics is needed.
Realizability of $\cJ$ and $\bcH$ after each time step requires a convex-invariant IMEX scheme.
Since the realizable moments form a convex set, it is possible to construct a realizability-preserving method using a convex-invariant IMEX scheme for two-moment neutrino transport.
In addition, the physics of neutrino transport in CCSNe requires the IMEX scheme to be diffusion-accurate.
